---
title: "Project 2: Summer Movie Ratings"
description: |
  Analysis of movies that contain the word summer in them
author: Felicia Akinde
date: September 27, 2024
format: html
css: alignment.css
about:
  template: solana
  links:
    - text: Summer Movies Data Source
      href: https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-07-30/readme.md
---

Project 2 task: Analyze data using: at least 3 str_*() functions, at least 3 regular expressions, at least 2 illustrative, well-labeled plots or tables, a description of what insights can be gained from your plots and tables, and a reference / documentation of the data source.


```{r message=FALSE, warning=FALSE}

library(tidyverse)

# imports data from tidytuesday
tuesdata <- tidytuesdayR::tt_load(2024, week = 31)

# renames data
summer_movie_genres <- tuesdata$summer_movie_genres
summer_movies <- tuesdata$summer_movies

```

```{r}
# creates tibble of summer movies that contain at least one digit
summer_movies_digi <- summer_movies |>
  filter(str_detect(simple_title, "\\d")) |>
  select(primary_title, year, genres, simple_title, average_rating)

# creates tibble of summer movies that include the specified pattern
summer_movies_of_digi <- summer_movies_digi |>
  filter(str_detect(simple_title, "summer (..) \\d{2,}"))

```

This is a table created from a data set that contains movies with the word "Summer" in them. This new table shows if they have the sequence "Summer" plus any two letters (most commonly of or in), plus a sequence of digits if those digits are more than double digits.

```{r}
# genres and average rating (plot)

movies_gen_rating <- summer_movies |>
  mutate(genres = str_split(genres, ",")) |>
  mutate(perc_over_5 = str_count(average_rating, "[5-9]\\.\\d+")) |>
  unnest(genres) |>
  filter(!is.na(genres)) |>
  group_by(genres) |>
  summarize(perc_over_5 = sum(perc_over_5)) |>
  select(genres, perc_over_5)

# plot
ggplot(movies_gen_rating, 
       aes(x = genres, y = perc_over_5)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Summer Movie Genre",
    y = "Amount of Ratings over 5.0",
    title = "Amount of over 5.0 Ratings per Summer Movie Genre",
  )
```
alt txt: Bar chart showing the amount of ratings over 5.0 for different summer movie genres. The x-axis represents movie genres, and the y-axis shows the proportion or count of ratings exceeding 5.0.


This is a graph relating genres of movies with "Summer" in the title, and the amount of 5.0+ ratings movies with that genre received. This graph shows that Drama, Comedy, and Romance all got the most ratings over 5.0. Film-Noir, Reality-TV, Talk-Show, and Western got the least ratings over 5.0.

Works Cited

Data sourced from TidyTuesday (linked in button) 

R for Data Science. (2024, July 30). TidyTuesday data: README file. GitHub. Retrieved December 6, 2024, from https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-07-30/readme.md

IMDb. (n.d.). Non-commercial datasets. IMDb Developer. Retrieved December 6, 2024, from https://developer.imdb.com/non-commercial-datasets/
