---
title: "Project 2: Summer Movie Ratings"
description: |
  Tbd
author: Felicia Akinde
date: September 27, 2024
format: html
css: alignment.css
about:
  template: solana
  links:
    - text: Summer Movies Data Source
      href: https://github.com/rfordatascience/tidytuesday/blob/master/data/2024/2024-07-30/readme.md
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)

tuesdata <- tidytuesdayR::tt_load(2024, week = 31)

summer_movie_genres <- tuesdata$summer_movie_genres
summer_movies <- tuesdata$summer_movies

```

```{r}
#
summer_movies_digi <- summer_movies |>
  filter(str_detect(simple_title, "\\d")) |>
  select(primary_title, year, genres, simple_title, average_rating)

summer_movies_of_digi <- summer_movies_digi |>
  filter(str_detect(simple_title, "summer (..) \\d{2,}"))

summer_movies_of_digi

```

This is a table created from a data set that contains movies with the word "Summer" in them. This new table shows if they have the sequence "Summer" plus any two letters (most commonly of or in), plus a sequence of digits if those digits are more than double digits.

```{r}
# genres and average rating (plot)

movies_gen_rating <- summer_movies |>
  mutate(genres = str_split(genres, ",")) |>
  mutate(perc_over_5 = str_count(average_rating, "[5-9]\\.\\d+")) |>
  unnest(genres) |>
  filter(!is.na(genres)) |>
  group_by(genres) |>
  summarize(perc_over_5 = sum(perc_over_5)) |>
  select(genres, perc_over_5)

ggplot(movies_gen_rating, 
       aes(x = genres, y = perc_over_5)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Summer Movie Genre",
    y = "Amount of Ratings over 5.0",
    title = "Amount of over 5.0 Ratings per Summer Movie Genre",
  )
```

This is a graph relating genres of movies with "Summer" in the title, and the amount of 5.0+ ratings movies with that genre received. This graph shows that Drama, Comedy, and Romance all got the most ratings over 5.0. Film-Noir, Reality-TV, Talk-Show, and Western got the least ratings over 5.0.

Works Cited

Data sourced from TidyTuesday (linked in button) and https://developer.imdb.com/non-commercial-datasets/
